---
title: "Wildlife Insights Data Exploration"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
editor_options: 
  chunk_output_type: console
---


```{r adjustable options, echo=F, include=F}
# Hello, before you click "Knit' above, please download your wildlife insights data, unzip it, and paste it into the 'raw data' folder. Thank you!

# Set your timezone  
tz <- "PET"

# Use a category to colour plots 
category <- "subproject_name"

# Define a palette for the colourscheme (from the metbrewer package)
col_pal <- "VanGogh2"

```



```{r non-adjustable options, echo=F, include=F}
# Hello, before you click "Knit' above, please download your wildlife insights data, unzip it, and paste it into the 'raw data' folder. Thank you!


# Load required packages 
list.of.packages <- c("leaflet", "dplyr", "kriging", "corrplot", "lubridate", "kableExtra", "rredlist","sf", "usedist", "ggplot2", "ggpubr", "googledrive", "purrr", "plotly", "googlesheets4", "MetBrewer")

# Check you have them and load them
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages, ,repos = "http://cran.us.r-project.org")

lapply(list.of.packages, require, character.only = TRUE)
```


```{r data setup, echo=F, include=F}
# Read in data frames
dep <- read.csv("raw_data/deployments.csv", header=T)
pro <- read.csv("raw_data/projects.csv", header=T)

# Check how many image files you have (WI splits csvs if they have >500,000 rows)
tmp <- list.files("raw_data/")
tmp <- tmp[substr(tmp,1,3)=="ima"]
# Load your data 
dat<- list()
i<- 1
for(i in 1:length(tmp))
{
  dat[[i]] <- read.csv(paste0("raw_data/", tmp[i]), header=T)
}
dat <- bind_rows(dat)

# Format the date objects
dat$timestamp <- strptime(dat$timestamp, "%Y-%m-%d %H:%M:%S", tz="America/Lima")
dep$start_date <- strptime(dep$start_date, "%Y-%m-%d", tz="UTC")
dep$end_date <- strptime(dep$end_date, "%Y-%m-%d", tz="UTC")

#Calculate days active
dep$days <- dep$end_date-dep$start_date


# Add deployment coverage
tmp <- st_as_sf(dep,coords=c("longitude", "latitude"),crs=4326)

```


# Deployments
To date we have `r length(unique(dep$deployment_id))` camera deployments from `r length(unique(paste(dep$longitude, dep$latitude)))` unique locations (based on X-Y coordinates). 

```{r, echo=F, include=F}
# Generate colours to display the category levels - R needs them as a factor
table(dep$subproject_name)
dep[,category] <- factor(dep[,category])

col.cat <- as.character(met.brewer(col_pal,n=length(levels(dep[,category]))))

dep$Cols <- col.cat[dep[,category]]
```

```{r map, echo=F, warning=F, message=F}

# Count the number of camera ststions
n.stat <- length(unique(dep$deployment_id))

m <- leaflet() %>%
  addProviderTiles(providers$Esri.WorldTopoMap, group="Base") %>%
  addProviderTiles(providers$Esri.WorldImagery, group="Satellite") %>%  # Add satellite data
  
  addCircleMarkers(lng= dep$longitude, lat= dep$latitude,
                   color= dep$Cols,
                   popup=paste( dep$deployment_id), radius=3) %>%
 addLegend("topleft", colors = col.cat,  labels = levels( dep[,category]),
    title = category,
    labFormat = labelFormat(prefix = "$"),
    opacity = 1
  ) %>%
  # Layers control
  addLayersControl(
    baseGroups = c("Satellite", "Base"),
    options = layersControlOptions(collapsed = FALSE)
  )
m

```


# Wildlife Insights Summary

```{r, echo=F}
tmp <- as.data.frame(table(dat$identified_by=="Computer vision"))
tmp$Type <- "Labelled"
tmp$Type[tmp$Var1==TRUE] <- "To label"

```

To date, `r Sys.Date()`, we have `r nrow(dat)` images in Wildlife Insights, of which `r round((tmp$Freq[tmp$Type=="Labelled"]/sum(tmp$Freq))*100, 1)`% are identified (labelled) by people:

```{r, echo=F}
par(mfrow=c(1,1))

f1 <- plot_ly(tmp,y=~tmp$Freq, x=tmp$Type,  type="bar", name=tmp$Type) %>% 
        layout(yaxis = list(title = 'Images'), barmode = 'group')

f1
```

The breakdown by labeler is as follows:

```{r, echo=F}
tmp2 <- as.data.frame(table(dat$identified_by[!dat$identified_by%in% c("Computer vision", "")]))

colnames(tmp2)[1] <- "Person"

f2 <- plot_ly(tmp2,y=~tmp2$Freq, x=tmp2$Person,  type="bar", name=tmp2$Person) %>% 
        layout(yaxis = list(title = 'Images labelled'), barmode = 'group')


f2

```

## Species labels

```{r, echo=F}
# Reduce to verfied data
vdat <- dat[!dat$identified_by %in% c("Computer vision", ""),]
# Remove blanks
vdat$sp <- paste(vdat$genus, vdat$species)
vdat <- vdat[vdat$is_blank==0,]
vdat <- vdat[vdat$species!="No CV Result",]
vdat <- vdat[vdat$common_name!="Blank",]

# Do all species with a latin name have a common name - YES use common name
#as.data.frame(table(paste(vdat$sp, vdat$common_name)))

# if the species name is blank it must be to family, order or class level
```

To date we have `r nrow(vdat)` image labels verified by human observers (after removing `r nrow(dat[dat$is_blank==1 | dat$species=="No CV Result" | dat$common_name =="Blank",])` blanks). 

We have `r length(unique(vdat$common_name))` unique classifications. Of these, `r length(unique(vdat[vdat$species!="",]$sp))` are to full species level, `r length(unique(vdat[vdat$species=="",]$genus))` to genus level, the rest are higher (family/order). 

Note you can use the slider bar at the bottom to control the the x-axis extent. 

```{r, echo=F}
layout(matrix(c(1,1,2), 1, 3, byrow = TRUE))
det.sum.total <- as.data.frame(count(vdat[vdat$is_blank==0,], common_name))
det.sum.total <- det.sum.total[order(det.sum.total$n),]


det.sum.total$order <-1:nrow(det.sum.total)

yform <- list(categoryorder = "array",
              categoryarray = det.sum.total$common_name,
              tickfont = list(size = 7))

xform <- list(rangeslider = list(type = "n"))

fig <- plot_ly(x = det.sum.total$n, y = det.sum.total$common_name, type = 'bar', orientation = 'h',
               height=2000)%>% 
 layout(yaxis = yform, xaxis=xform)

fig

```

# Diel time
For any species detected more than 5 times, we will plot when they were detected:

```{r, echo=F, warning=F, message=F}
vdat$hours <- hour(vdat$timestamp) + minute(vdat$timestamp)/60 + second(vdat$timestamp)/(60*60)

# Susbet to more than 100 captures
tmp <- det.sum.total[det.sum.total$n>100,]

yform <- list(categoryorder = "array",
              categoryarray = tmp$common_name,
              tickfont = list(size = 7))
tmp2 <- vdat[vdat$common_name %in% yform$categoryarray,]

fig <- plot_ly(x = tmp2$hours, y = tmp2$common_name,type="scatter",
               height=1000, text=tmp2$deployment_id, hoverinfo='text',
               mode   = 'markers',
               marker = list(size = 5,
                             color = 'rgba(50, 100, 255, .2)',
                             line = list(color = 'rgba(0, 0, 0, 0)',
                                         width = 0))) %>% 
              layout(yaxis = yform)

fig

```


# Detection check

We also need to check if all of the deployments in Wildlife Insights have associated image data.

Of the deployments currently in Wildlife Insights, **`r paste0(round( (length(dep$deployment_id[dep$deployment_id %in% dat$deployment_id])/nrow(dep))*100,2), "%")`** of deployments have associated image data.

Check we don't have duplicates: `r length(dep$deployment_id[duplicated(dep$deployment_id)==TRUE])`

```{r, echo=F, eval=F}
#If that number is >0, what observations does it relate to?
dep[dep$deployment_id %in% dep$deployment_id[duplicated(dep$deployment_id)==TRUE],]
```

*Note - typical workflows should mean that deployments ALWAYS have image data (they are created and uploaded at the same time). If the percentage is <100% then there is a problem*

## Check for date-time camera malfunctions

Sometime cameras get time and date manfunctions. One symptom of this would be images before the first deployment date. Do we have that in this instance? 

We have `r nrow(dat[dat$timestamp < min(dep$start_date),])` images from  `r length(unique(dat$deployment_id[dat$timestamp < min(dep$start_date)]))` deployments which were taken before the start date. The deployments and the year the images are from are:

```{r, echo=F}
unique(paste(dat$deployment_id[dat$timestamp < min(dep$start_date)], substr(dat$timestamp[dat$timestamp < min(dep$start_date)],1,4)))
```

## Detection plots

The following series of plots allows you to check if your cameras are collecting photos when you thought they were. The black bars denote camera activity, the vertical lines denote detections: 

```{r, echo=F}
#We can also check to see if any deployments need truncating - for example if the camera ceased to be active:
dat$sp <- paste(dat$genus, dat$species)


# Plot the start and end date for each deployment
#fig <- plot_ly(x = dep$start_date[1:5], y = dep$deployment_id[1:5],type="scatter",
#               height=2000)
plot.order <- dep$deployment_id[order(dep$deployment_id)]
# Remove duplicates
plot.order<- plot.order[duplicated(plot.order)==F]
#fig
tmp4 <- data.frame("deployment_id"=plot.order, "Plot.grp"=ceiling(1:length(plot.order)/20))

eff <- left_join(tmp4,dep,by="deployment_id")
eff <- eff[duplicated(eff)==FALSE,]


i<- 15
j <- 1
for(j in 1:length(unique(eff$Plot.grp)))
{
    layout(matrix(c(1,1,1,2), 1, 4, byrow = TRUE))
    par(mar=c(2,12,1,1))
    #str(eff)
    plot(c(min(eff$start_date, na.rm=T), max(eff$end_date, na.rm=T)),      c(1,length(unique(eff$deployment_id[eff$Plot.grp==j]))), las=1, ylab="", xlab="", type="n", yaxt="n")
    
    axis(2, at= length(unique(eff$deployment_id[eff$Plot.grp==j])):1, labels= unique(eff$deployment_id[eff$Plot.grp==j]), las=1, cex.axis=1)
    #mtext("Camera Deployment ID", 2, 4)
    # Make lines for each of the cameras
    for(i in length(eff$deployment_id[eff$Plot.grp==j]):1)
    {
      abline(h=i, col=rgb(0,0,0,0.1))
      tmp <- eff[eff$deployment_id==rev(eff$deployment_id[eff$Plot.grp==j])[i],]
      
      tmp2 <- dat[dat$deployment_id==tmp$deployment_id[1],]
      #tmp2 <- left_join(tmp2, tmp3, by = "sp")
      points(tmp2$timestamp, rep(i,nrow(tmp2)), pch="|", col= tmp2$is_blank+2)
    
      for(k in 1:nrow(tmp))
        {
          lines(c(tmp$start_date[k],
                           tmp$end_date[k]),
                c(i,i), lwd=2)
        }
      }
    par(mar=c(0,0,1,0))
    plot.new()
    legend("topleft", legend=c("Animal", "Blank"), fill=c(2,3), xpd=TRUE, cex=1.1 )

}


```



# Full species list

A full list of labelled species in alphabetic order:

```{r, echo=F}
#head(vdat)
tmp <- vdat[, c("class","order","family", "genus","species")]

tmp <- tmp[duplicated(tmp)==F,]
tmp <- tmp[ with(tmp, order(class,order,family, genus,species)),]
row.names(tmp) <- NULL

tmp %>%
  kbl() %>%
  kable_styling(full_width = T) %>%
  #column_spec(1, bold = T, border_right = T )%>%
  scroll_box( height = "800px")




```

# Meta data

Summary information for the project:

```{r, echo=F}
t(pro) %>%
  kbl() %>%
  kable_styling(full_width = T) %>%
  column_spec(1, bold = T, border_right = T)


```


```{r, echo=F}
# Write updated files
dir.create("data_processed")
write.csv(eff,"data_processed/Effort.csv")
```


